Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations

### 2024-03-16 notebooks/events/index.qmd
update

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_orbit.png) ![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_HEE.png)

>>>>

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_orbit.png)

![](https://psp-gateway.jhuapl.edu/website/SciencePlanning/Encounters/E8_HEE.png)

### 2024-03-16 presentations/sparthb.qmd
update

---
title: "Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
  - name: Chen Shi
  - name: Zesen Huang
# format: revealjs
format:
  pptx:
    slide-level: 2
  html: default
  revealjs:
    output-file: 'sparthb.revealjs.html'
slide-number: true
---

@@

@@

Studying the radial distribution of occurrence rate, as well as the properties of solar wind discontinuities may help answer the following questions:

@@

## Evolution with radial distance

@@

![](../notebooks/events/images/paste-1.png)
:::

@@

<!-- ![Normalized solar wind density at the ecliptic and the IMF lines](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_den1-psp-x15wifs3z1fo_med_201903-201904.mp4) -->

<!-- ![Radial solar wind velocity at the ecliptic on the global and detail regions together with the temporal profile at spacecraft](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_vel2e1-psp-x15ifs3z1fr_med_201903-201904.mp4) -->

@@

::: {.column}
![ACE observation (continued)](../images/paste-17.png)
:::

@@

Solar wind discontinuities examples observed by PSP and ARTEMIS fitted by a logistic function in LMN coordinates.

In general discontinuities observed by PSP is much shorter (1~5 secs) with larger current density.

@@

@@

Every discontinuity's L component is fitted by a tanh (logistic) function.

@@

$\Delta v_l$

@@

@@

@@

@@

>>>>

---
title: "Solar Wind Discontinuities: Parker Solar Probe vs ARTEMIS Observations"
subtitle: SPARTHB
author:
  - name: Zijin Zhang
  - name: Anton V. Artemyev
  - name: Vassilis Angelopoulos
  - name: Chen Shi
  - name: Zesen Huang
# format: revealjs
format:
  html: default
  pptx:
    slide-level: 2
  revealjs:
    output-file: 'sparthb.revealjs.html'
slide-number: true
---

@@

- Import element of solar wind turbulences

- Very kinetic structure: interact with energetic part of the solar wind ions

@@

Studying the radial evolution of solar wind discontinuities from synergistic observations of PSP and Earth-orbiting missions (ARTEMIS, Wind) during aligned intervals.

@@

## In-situ synergistic observations

**Similar** type of solar wind or **Same** solar solar wind

How to define alignment?

- Trajectory & Orbits (theory): ballistic approximation of Parker spiral

- Plasma properties (observation):

  - Invariants: plasma composition, ionization states, mass flux $n u r^2$

  - Correlation: magnetic field / plasma velocity

- Simulation: Trace plasma parcels evolution

<!-- sensing solar wind emitted from the same region on the solar surface, with the caveat that dynamical processes at the chromosphere–corona interface have not dramatically altered the solar footpoint within the time corresponding to the radial propagation from the measurement of one spacecraft to that of the other. -->

@@

![Heliographic frame (corotating with Sun)](../notebooks/events/images/paste-1.png)
:::

@@

::: {.content-visible when-format="html"}

![Normalized solar wind density at the ecliptic and the IMF lines](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_den1-psp-x15wifs3z1fo_med_201903-201904.mp4)

![Radial solar wind velocity at the ecliptic on the global and detail regions together with the temporal profile at spacecraft](../images/enlil/tuz-a7b1-d4t05x1-d200t1-donki_vel2e1-psp-x15ifs3z1fr_med_201903-201904.mp4)
:::

@@

::: {.column}
![ACE observation (continued)](../images/paste-17.png)
:::
::::

---

Time resolution (from high to low): SPC, SPI, SQTN, QTN

Quality (fluctuating, from high to low): SQTN, QTN, SPC/SPI

![Different Density Product ](../figures/examples/psp_density_comparison.png)

@@

In general discontinuities observed by PSP are with

- shorter duration (1~5 secs)

  compared with 5-10 secs

- large current density (100-1000 nA/m^2)

  compared with 1-10 nA/m^2

@@

![](../figures/examples/detection_liu2022.png)

@liuMagneticDiscontinuitiesSolar2022

Compared with adjacent intervals without sampling

@@

Every discontinuity's $l$ component is fitted by a $\tanh$ (logistic) function.

Combined with plasma data ($n$, $v$, $T$)

@@

---

If we expect the discontinuities to be generated locally, then the thickness of the discontinuities should be related to local plasma parameters:

- Normalize thickness by $d_i$ (ion inertial length) or $r_i$ (ion gyroradius)
- Normalize current density by $J_A$ (Alfven current density)

@@

![](../figures/enc2/thickness_current.spi_n_spi.png)

![](../figures/enc2/density_distribution.spi_n_spi.png)

K1: Normalized discontinuities properties does not vary much with radial distance

---

![](../figures/enc2/thickness_current.spc_n_spc.png)

![](../figures/enc2/density_distribution.spc_n_spc.png)

K2: $Δv_l$,l demonstrates a strong correlation with $Δv_{A}$, albeit being consistently smaller.

@@

For rotational discontinuities, the plasma velocity jump across them: the plasma ﬂow velocity $v_l$ (the solar wind velocity projected onto $l$) changes due to changes in $B_l$, $Δv_l=±Δv_A$ with $v_A= B_l / \sqrt{4 n m}$ [@hudsonDiscontinuitiesAnisotropicPlasma1970].

![](../figures/enc2/dvl.spi_n_spi.png)

@@

The equation for $Δv_A$ includes a factor depending on the anisotropy of the plasma.

![](../figures/enc2/anisotropy.spi_n_spi.png)

Pressure anisotropy $Λ = \mu (P_∥ - P_⊥) / B^2$.

@@

- The properties of the discontinuities is related to the local plasma parameters
  - Thickness => ion inertial length
  - Current density => Alfven velocity (current density)
- Normalized thickness and current density of discontinuities remain constant with radial distance
- Anisotropy of the plasma is expected to be larger near the sun to explain the observed speed change ratio $Δv_i/Δv_A$.

---

## References
